Semantic Novelty Detection with Graphical Models

### 2011-05-05 semantic-nd/article.tex
advances on explaining dora and graphs

% Additional packages
\usepackage[utf8]{inputenc}
\usepackage{tikz}

@@

% first the title is needed
\title{Semantic Novelty Detection with Graphical Models
on Dora Architecture}

@@

\section{Related Work}
Although this paper deals with novelty detection on the robotics area, it does so using very
standard concepts and techniques such as semantic data and graphical models.
Those are often found on areas related with information retrieval.

Interesting examples are works in automatic image annotation using an hidden concept layer between visual features and text information\cite{multi-model-image-retrieval}.

\section{Dora Architecture Overview}
\begin{itemize}
\item Introduce Dora architecture aspects important for the paper
\item Explain what we mean by semantic data and how its captured by Dora
\item Introduce graphical models and how we use them (including notation on the paper)
\end{itemize}

Dora is composed by several layers, though for the sake of this paper only the conceptual layer is important.

\subsection{Semantic Data}
Dora has a set of machine learning tools that are used to sense the environment around it.
The output of those describe a certain high-level property or information that we refer as \emph{semantic data}.
The following types of probabilistic information are extracted by Dora and used on this paper:
\begin{description}
\item[Object detection] is performed by running several pre-trained object detectors on the visual input. The system becomes
aware of the presence of objects such as book, cereal box, computer, robot, stapler, toilet paper.

\item[Room Size and Shape] are extracted by using the 2D laser scans. The system categorizes each room in terms of its size (large, medium, small) and shape (rectangular and elongated).

\item[Room Appearance] is categorized using CRFH and a pre-trained set of 7 different models.

\item[Doorway detection] is used to segment the low-level space into rooms and map connectivity between them.
\end{description}

\subsection{Graphical Models}
\begin{itemize}
\item Explain Dora use of Graphical Models
\end{itemize}

Dora uses \emph{chain graphs} as modelling tool. Those can be converted to \emph{factor graphs}\cite{factor}.

@@

The edges and factors between the several variables can be seen as a kind of filter to
the properties of the system.

Graphical models provide a generative model for a distributions of variables.
Those are mainly interesting since they allows to calculate the probability of any set of variables $x$

As they can be seen as a method to represent distributions of variables. From now on we shall use the notation $P_G(x)$ to denote the probability that the distribution described by a graphical model $G$ generates the set $x$ of variables.

\section{Novelty Detection}
\begin{itemize}
\item Explain 
\item Approach with a threshold function
\item Need of a normalizing factor for a threshold with dynamic functions
\item Usage of conditional density
\item Assuming of every output is equally likely
\item Using unlabelled data
\item What happens if the conditional and unconditional graphs differ only by the addition of a node?
\item How to implement that in the case a graph is a tree
\item Novelty detection on any variable of the graphical model!
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{4}

\bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular
Subsequences. J. Mol. Biol. 147, 195--197 (1981)

\bibitem{lncschap} May, P., Ehrlich, H.C., Steinke, T.: ZIB Structure Prediction Pipeline:
Composing a Complex Biological Workflow through Web Services. In: Nagel,
W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006. LNCS, vol. 4128,
pp. 1148--1158. Springer, Heidelberg (2006)

\bibitem{book} Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing
Infrastructure. Morgan Kaufmann, San Francisco (1999)

\bibitem{proceeding1} Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid
Information Services for Distributed Resource Sharing. In: 10th IEEE
International Symposium on High Performance Distributed Computing, pp.
181--184. IEEE Press, New York (2001)

\bibitem{proceeding2} Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The Physiology of the
Grid: an Open Grid Services Architecture for Distributed Systems
Integration. Technical report, Global Grid Forum (2002)

\bibitem{url} National Center for Biotechnology Information, \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}

>>>>

% Additional packages
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{tikz}
\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
\hypersetup{%
   plainpages=false, 
   pdfpagelayout=SinglePage,
   bookmarksopen=false,
   bookmarksnumbered=true,
   breaklinks=true,
   linktocpage,
   colorlinks=true,
   linkcolor=blue,
   urlcolor=blue,
   citecolor=blue,
   anchorcolor=green
}

@@

% first the title is needed
\title{Semantic Novelty Detection with Graphical Models}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\cite{quattoni2009recognizing} showed that most scene recognition models work poorly in indoor
scenes when compared to outdoor scenes results.
Since the properties that characterize rooms changes conforming its category. Namely corridors are
well described by global properties and bookstores are well described by the presence of specific objects (books).
Their work shows a multi-modal approach is expected to yield better performance by mixing several sources of information.

\cite{galindo2005multi} defines a bidirectional relation between object and room category, where object defines a room category and a room category provides information on where objects may be found.

Probabilistic representations are used in several localised functions in robots operating in the real-world~\cite{gross2009toomas,maierprobabilistic}. And some employ, up to some extent, a probabilistic representation across some subsystems~\cite{kraft2008exploration}.
\cite{vasudevan2008bayesian} performed room categorization through Bayesian reasoning about the presence of objects but did not included observations models (perception is considered deterministic).
And \cite{boutell2006factor} have studied outdoor scene classification using \emph{factor graphs} and modelling spatial relations between objects in the scene to extract better knowledge from semantic (high-level) features.

Its expected that using a unified probabilistic model from the whole system, such as \cite{pronobis2011exploiting}, more information can be reused to correctly predict a given random variable.

Although this paper deals with novelty detection on the robotics area, it does so using very
standard concepts and techniques such as semantic data and graphical models.
Those are often found on areas related with information retrieval.
Interesting examples are works in automatic image annotation using an hidden concept layer between visual features and text information\cite{zhang2005probabilistic}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dora Architecture Overview}
\begin{itemize}
\item Introduce Dora architecture aspects important for the paper
\item Explain what we mean by semantic data and how its captured by Dora
\item Introduce graphical models and how we use them (including notation on the paper)
\end{itemize}

Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora.

From its architecture only the \emph{conceptual layer} is of interest to this article.
Its role is to aggregate the following semantic information coming from other layers:

\begin{description}
 \item[Doorway detection] is used to segment the low-level space into rooms and map connectivity between them.
 \item[Room size and shape] are classified by using 2D laser scans data and are associated as properties of a given room. The system utilizes pre-trained set of classifiers to label size as (large, medium, small) and shape as (rectangular or elongated).
 \item[Object detection] is performed using the visual input in order to detect . The system keeps track of the number of object and types seen per each room. Objects are once again detected by running a pre-trained set of detectors for: book, cereal box, computer, robot, stapler, toilet paper.
 \item[Room appearance] is categorized from the visual input by using CRFH and a pre-trained set of 7 different models.
\end{description}

With the extracted information and conceptual knowledge the conceptual layer creates a structured probabilistic representation
in order to model all known variables and their relations.
As so, in order to represent the knowledge the layer builds a \emph{chain graph}:
a probabilistic graphical model that merges both Bayesian Networks and Random Markov Fields.

The use of graphical models to describe distributions of variables has useful properties.
The edges between the variables can be seen as a kind of filter to the properties of the system.
Properties those that we expect to capture by the conceptual knowledge.
At the same time they are a generative models and therefore allow to calculate the probability
on any given subset of variables on the graph allowing the system to work even when some information is missing.

\begin{figure}[h]
\centering
\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge in order to
         create a structured probabilistic representation of the world.}
\end{figure}

\subsection{Factor Graphs}
Although the conceptual layer works with \emph{chain graph}, those can be converted into \emph{factor graphs}.
We opted to use those as describing the ideas on this paper as they tend to be easier to understand.

A \emph{factor graph} is a bipartite graph connecting nodes representing random variables and factors.
Each factor represents a function dependent only on the variables to where its connected.
As so given factor graph $G$ can be seen as description of a distribution function over a set of variables obtained by
the product of all the factors. This manipulation allows to use algorithms such as the Sum-Product to efficiently calculate
probabilities on any given subset $x$ of those variables by exploiting conditional independency between variables.
We will use the notation $P_G(x)$ to represent that.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Novelty Detection}
\begin{itemize}
\item Explain 
\item Approach with a threshold function
\item Need of a normalizing factor for a threshold with dynamic functions
\item Usage of conditional density
\item Assuming of every output is equally likely
\item Using unlabelled data
\item What happens if the conditional and unconditional graphs differ only by the addition of a node?
\item How to implement that in the case a graph is a tree
\item Novelty detection on any variable of the graphical model!
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plainnat}
\bibliography{refs}
